\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-05-20 paper.qmd
Added Scurve model in data space screen shots

#### Generating edges to indicate neighbours and removing long edges

Delaunay triangulation is used to connect neighbouring centroids, which is needed to preserve neighbourhood information when the model is lifted into \pD{}.

When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in 2D space, which is a triangular mesh (@fig-NLDR-scurve (d)). It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.

@@

<!--langevitour with pD model for S-curve-->

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 190, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  theme(axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 5),
        axis.title.y = element_text(size = 5, angle = 90))
```

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices. What is the effective $b$ to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE. Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE. Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
#| label: fig-mse-scurve-b
#| fig-pos: H
#| out-height: 20%

>>>>

#### Generating edges to indicate neighbors and removing long edges

Delaunay triangulation is used to connect neighboring centroids, which is needed to preserve neighborhood information when the model is lifted into \pD{}.

When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in \gD{} space, which is a triangular mesh (@fig-NLDR-scurve (d)). It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.

@@

<!--langevitour with pD model for S-curve-->
::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
![](figures/scurve/sc_1.png){#fig-scurve_sc1}

![](figures/scurve/sc_2.png){#fig-scurve_sc2}

![](figures/scurve/sc_3.png){#fig-scurve_sc3}

Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/G1m4q9k--v4>).
:::

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 190, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  theme(axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7, angle = 90))
```

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices. What is the effective $b$ to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE. Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE. Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
#| label: fig-mse-scurve-b
#| fig-pos: H
#| out-height: 50%

### 2024-05-20 paper.tex
Added Scurve model in data space screen shots

\subsubsection{Generating edges to indicate neighbours and removing long
edges}\label{generating-edges-to-indicate-neighbours-and-removing-long-edges}

Delaunay triangulation is used to connect neighbouring centroids, which
is needed to preserve neighbourhood information when the model is lifted
into \pD{}.

When \(k = 2\) Delaunay triangulation on \(C^{(2)}\) generates the model
in 2D space, which is a triangular mesh (Figure~\ref{fig-NLDR-scurve}
(d)). It generates convex hulls of \(C^{(2)}\) such that the
circumcircle of every triangle in the triangulation contains no other
points from \(C^{(2)}\).

@@

@@

\includegraphics[width=1\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}

>>>>

\subsubsection{Generating edges to indicate neighbors and removing long
edges}\label{generating-edges-to-indicate-neighbors-and-removing-long-edges}

Delaunay triangulation is used to connect neighboring centroids, which
is needed to preserve neighborhood information when the model is lifted
into \pD{}.

When \(k = 2\) Delaunay triangulation on \(C^{(2)}\) generates the model
in \gD{} space, which is a triangular mesh (Figure~\ref{fig-NLDR-scurve}
(d)). It generates convex hulls of \(C^{(2)}\) such that the
circumcircle of every triangle in the triangulation contains no other
points from \(C^{(2)}\).

@@

\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_1.png}

\subcaption{\label{fig-scurve_sc1}}

\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_2.png}

\subcaption{\label{fig-scurve_sc2}}

\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_3.png}

\subcaption{\label{fig-scurve_sc3}}

\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
of the S-curve, shows the model-in-data space, a video of the tour
animation is available at (\url{https://youtu.be/G1m4q9k--v4}).}

@@

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
